---
title: "Census Map"
author: "Giovanna Lino"
---

## **Introduction**

In this walkthrough, I will demonstrate how to create an interactive county census map of a chosen state using R. Specifically, I will be using the **`tidycensus`** package to download census data from the American Community Survey and the **`tmap`** and **`mapview`** packages to create the interactive map.

I have chosen the state of Texas and will be mapping the percentage of individuals over the age of 25 with a bachelor's degree or higher.

## **Step 1: Install Packages**

First, make sure you have the necessary packages installed and loaded. In this case I have used the following packages:

```{r}

install.packages("tidycensus")
install.packages("tmap")
install.packages("mapview")

# Load packages
library(tidycensus)
library(tmap)
library(mapview)
library(dplyr)

```

## **Step 2: Set up Census API Key**

To download data from the American Community Survey using the **`tidycensus`** package, you need to set up a Census API key.

With the API key,set it up in R using the following code:

```{r}
# Census API key
#Retrived from the API Link on their page 
census_api_key("75fa32dac9122cf2c43fb6e5d0ed1e7ead626185")

```

## **Step 3: Download Census Data**

Download the census data for Texas using the **`tidycensus`** package. Specifically, we will be downloading the percentage of individuals over the age of 25 with a bachelor's degree or higher.

```{r}

# Download census data
tx_census <- get_acs(
  geography = "county",
  variables = "B15003_022",
  state = "TX",
  survey = "acs5",
  year = 2019)

```

## **Step 4: Join Census Data with Spatial Data**

Now, join the census data with spatial data for Texas counties. We will be using the **`tigris`** package to get the county shapefiles and join them with the census data using the **`left_join`** function.

```{r}

# Load tigris package
library(tigris)

# Texas county shapefiles
tx_counties <- counties(state = "TX", cb = TRUE)

# Join census data with spatial data
tx_map_data <- left_join(tx_counties, tx_census, by = c("NAME" = "NAME"))
```

## **Step 5: Create Choropleth Map**

Create a choropleth map of our data using the **`tmap`** package.

```{r}

tm_shape(tx_map_data) +
  tm_polygons(variables = "B15003_022",
              col = "grey", 
              border.col = "black",
              style = "quantile",
              palette = "Blues",
              title = "Percentage of Individuals with more than 25 with Bachelor's Degree or Higher") +
  tm_layout(main.title = "Texas Counties",
            legend.outside = TRUE)

```
